---
title: "How to Work with Data from a Subsystem"
author: "Pri Oberoi"
date: "August 9, 2016"
output: html_document
---

Data from the Net Zero house includes measurements from various instruments around the house in one minute increments. Each variable, also called channels in  NIST documentation, provides the readings from one instrument, and the instruments are grouped into subsystems like HVAC, Electrical, DHW.

This tutorial walks through loading data from a particular subsystem.

## Get the data you need

If you want to analyze net-zero data for specific subsystems, rather than data for the whole house, you can download the data from the Net Zero website and import it into R, or you can use the download link directly.

For example, if you wanted to analyze the channels (like the heat pump air flow rate or the total power used by the heat pump) in the HVAC subsystem, you could pull data for that subsystem:

```{r, warning=FALSE}
# load libraries
library(ggplot2) # plotting library
library(DT) # interface to the JavaScript library DataTables

# get data
hvac <- read.csv("https://s3.amazonaws.com/net-zero/2015-data-files/HVAC.csv", header=TRUE, na.strings=c("NA", "NULL", "", " "))

```

Here are the variables included in the HVAC subsystem data. The nomenclature includes the name of the Sub-System and the name of the channel as SubSystem_SpecificChannel.

```{r, warning=FALSE}
datatable(data.frame(channels = names(hvac)), options = list(pageLength = 20))
```

## Working with the data

Once you have your data, you can visualize the Supply and Return air temperatures by season. 

First, convert TimeStampGPSTime into a date/time format using strptime():
```{r, warning=FALSE}
# converts TimeStampGPSTime variable into date/time format (POSIXlt)
hvac$TimeStampGPSTime <- strptime(hvac$TimeStampGPSTime, "%Y-%m-%d %H:%M:%S")

# here is a sample of what that looks like: 
hvac$TimeStampGPSTime[1:10]
```

Now we can use the TimeStampGPSTime variable to create a new categorical 'season' variable:
```{r, warning=FALSE}
# create a variable for seasons
hvac$season <- "summer"
hvac$season[hvac$TimeStampGPSTime < '2015-03-20 12:30:00'] <- "spring"
hvac$season[hvac$TimeStampGPSTime > '2015-09-22 10:21:00' & hvac$TimeStampGPSTime < '2015-12-21 05:44:00'] <- "fall"
hvac$season[hvac$TimeStampGPSTime > '2015-12-21 05:44:00'] <- "winter"
hvac$season <- as.factor(hvac$season)

# now we can do things like group readings by season
ggplot(hvac, aes(x = season, y = HVAC_HVACTempSupplyAir, fill = season)) +
  geom_violin(scale = "area") +
  ggtitle("HVAC Air Supply Temperature by Season")
```

Additionally, we can compare readings between channels in the subsystem; this compares the air temperature in the HVAC supply air with the retun air:
```{r, warning=FALSE}
# Visualize supply and return air temperature  
ggplot(hvac, aes(x = TimeStampCount)) +
  geom_point(aes(y = HVAC_HVACTempSupplyAir), colour = "blue", alpha = 0.10, size = 1) +
  geom_point(aes(y = HVAC_HVACTempReturnAir), colour = "black", alpha = 0.10, size = 1) +
  xlab("Time") +
  facet_wrap( ~ season, ncol = 2) +
  ggtitle("Seasonal Supply and Return Air Temperatures in the HVAC Subsystem")
```

