---
title: "How to Work with Data from a Subsystem"
author: "Pri Oberoi"
date: "August 9, 2016"
output: html_document
---

## Get the data you need

If you want to analyze net-zero data for specific subsystems, rather than data for the whole house, you can download the data from the Net Zero website and import it into R, or you can use the download link directly.

For example, if you wanted to analyze the channels (like the heat pump air flow rate and/or the total power used by the heat pump) in the HVAC subsystem, you could pull data for that subsystem:

```{r}
# load libraries
library(ggplot2) # plotting library
library(DT) # interface to the JavaScript library DataTables

# get data
hvac <- read.csv("https://s3.amazonaws.com/net-zero/2015-data-files/HVAC.csv", header=TRUE, na.strings=c("NA", "NULL", "", " "))

```

Here are the variables included in the HVAC subsystem data:

```{r}
datatable(data.frame(channels = names(hvac), minValue = apply(hvac, 2, min), maxValue = apply(hvac, 2, max), nullCount = apply(hvac, 2, function(x) sum(is.na(x)))), options = list(pageLength = 20))
```

## Visualize the data

Once you have your data, you can visualize the Supply and Return air temperatures by season. 

```{r}

# converts TimeStampGPSTime variable into date/time format (POSIXlt)
hvac$TimeStampGPSTime <- strptime(hvac$TimeStampGPSTime, "%Y-%m-%d %H:%M:%S")

# create a variable for seasons
hvac$season <- "summer"
hvac$season[hvac$TimeStampGPSTime < '2015-03-20 12:30:00'] <- "spring"
hvac$season[hvac$TimeStampGPSTime > '2015-09-22 10:21:00' & hvac$TimeStampGPSTime < '2015-12-21 05:44:00'] <- "fall"
hvac$season[hvac$TimeStampGPSTime > '2015-12-21 05:44:00'] <- "winter"
hvac$season <- as.factor(hvac$season)

# visualize air temperature  
ggplot(hvac, aes(x = TimeStampCount)) +
  geom_point(aes(y = HVAC_HVACTempSupplyAir), colour = "blue", alpha = 0.10, size = 1) +
  geom_point(aes(y = HVAC_HVACTempReturnAir), colour = "black", alpha = 0.10, size = 1) +
  xlab("Time") +
  facet_wrap( ~ season, ncol = 2) 

```

