---
title: "Analyzing Aggregated Readings"
output: html_document
---

Channels provide data at the instrumentation-level, however it is possible to analyze aggregate readings at the subsystem level.

This tutorial walks through analyzing aggregate/total instantaneous readings for channels in the Electrical subsystem

## Get the data you need

First, we can use the download link from the Data page to import the data for the Electrical Subsystem:

```{r, warning=FALSE}
# load libraries
library(ggplot2) # plotting library
library(reshape2) # library for reshaping data easily

# get data
#elec <- read.csv("https://s3.amazonaws.com/nist-netzero/2015-data-files/Elec.csv", header=TRUE, na.strings=c("NA", "NULL", "", " "))
# delete the line below and uncomment the line above
elec <- read.csv("/Users/prioberoi/Documents/nist/netZero/data/processed/Elec.csv", header=TRUE, na.strings=c("NA", "NULL", "", " "))

```

## Working with the data

There are 158 channels in the Electrical subsystem, however it is possible to aggregate those instantaneous power consumption readings.

```{r, warning=FALSE}
# Row and column count for Electrical channel data
dim(elec)

# create an object that only has instantaneous power readings
elec_power <- elec[,c("Timestamp", names(elec)[grep("_Power", names(elec))])]

# using the data dictionary or metadata files, this object could be narrowed down to power consumption in one room or for a specific appliance
# the 'total' variable below is a sum of all electrical power consumption channels
elec_power$total <- rowSums(elec_power[,2:ncol(elec_power)])
ggplot(elec_power[,c("Timestamp", "total")], aes(x = Timestamp, y = total)) +
  geom_point(aes(colour = strftime(elec_power$Timestamp, format = "%m")), alpha = 0.4) +
  geom_smooth() +
  theme(legend.position="none") +
  ggtitle("Total Instantaneous Power Consumption Readings")

```

